---
title: "Gov 50 Final Project"
author: Shivali Korgaonkar
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Milestone 2

Does a religious background make an LGBT individual less likely to come out? In this study, I will analyze how different religious backgrounds impact gay men, lesbians, bisexuals, and transgender individuals in coming out. I hypothesize that being religious makes people less comfortable coming out to family and friends. I expect that it is easier for people to come out when religion doesn’t play a big role in their life. My sample is composed of 1,197 survey results from gay men, lesbians, bisexuals, and transgender respondents who are 18 years or older. There is diversity in their religious backgrounds, as well as their status of discretion to the public on their sexuality and/or gender identity. The explanatory variable of interest is the religious status of the respondent, meaning their degree of religiousness, and if applicable, their religion. This is measured with two variables. First, =1 is Protestant, = 2 is Roman Catholic, =3 is Agnostic/Atheist, =4 is Nothing in Particular, =5 is Christian, and =6 is Other. Second, the importance of religion is measured as =1 is very important, =2 is somewhat important, =3 is not too important, and =4 is not at all important. The outcome variable is whether the important people in the individual’s life are aware of their sexuality and/or gender identity. The variable is measured as =1 means almost all know, =2 means some of them, =3 means a few, and =4 means none. The data comes from the KnowledgePanel research group in 2013. If the data shows that individuals with a low importance in religion have come out to almost all their loved ones, my hypothesis would be supported. If individuals with a heavy emphasis on religion come out at higher or equal levels as non-religious individuals, my hypothesis would be disproven. The additional variable of which religion they identify as can provide nuance in understanding which religions provide a more/less welcoming view towards the LGBT community coming out. When I run a regression on coming out status on non-religious individuals, a positive, significant coefficient would support my claim. 

```{r setup}

library(haven)
library(readr)

library(haven)
lgbt <- read_sav("pew2013lgbtpublicdatarelease (1)/pew2013lgbtpublicdatarelease.sav")

library(tidyverse)
library(ggplot2)
library(broom)
library(dplyr)
```


```{r}

lgbt1 <- lgbt |>
  select(NEWCASEID, LGBT, Q31, Q42, Q49, Q66_A:Q67, religNEW, IMP) |>
  mutate(LGBT = case_when(
    LGBT == 1 ~ "Lesbian",
    LGBT == 2 ~ "Gay",
    LGBT == 3 ~ "Bisexual",
    LGBT == 4 ~ "Transgender",
    LGBT == 5 ~ "Non-LGBT"
  )) |>
  mutate(LGBTQ_importance = case_when(
    Q31 == 1 ~ "Extremely Important",
    Q31 == 2 ~ "Very Important",
    Q31 == 3 ~ "Somewhat Important",
    Q31 == 4 ~ "Not too important",
    Q31 == 5 ~ "Not at all important"
  )) |>
  select(-Q31) |>
  rename("ComeOutAge" = "Q42") |>
  mutate(Important_People = case_when(
    Q49 == 1 ~ "All or most of them",
    Q49 == 2 ~ "Some of them",
    Q49 == 3 ~ "Only a few of them",
    Q49 == 4 ~ "None of them"
  )) |>
  rename("Evangelical" = "Q66_A",
         "Catholic" = "Q66_B",
         "Jewish" = "Q66_C",
         "Muslim" = "Q66_D",
         "Mormon" = "Q66_E",
         "Non-Evangelical Protestant" = "Q66_F") |>
  mutate(Religious_Conflict = if_else(Q67 == 1, 
                                      "Conflict",
                                      "No Conflict")) |>
  select(-Q67) |>
  mutate(Religion = case_when(
    religNEW == 1 ~ "Protestant",
    religNEW == 2 ~ "Roman Catholic",
    religNEW == 3 ~ "Agnostic/Atheist",
    religNEW == 4 ~ "Nothing",
    religNEW == 5 ~ "Christian",
    religNEW == 6 ~ "Other"
  )) |>
  mutate(Religion_Importance = case_when(
    IMP == 1 ~ "Very important",
    IMP == 2 ~ "Somewhat important",
    IMP == 3 ~ "Not too important",
    IMP == 4 ~ "Not at all important"
  )) |>
  filter(!Evangelical == -1)


```

```{r}
## 1- Friendly, 2- Neutral, 3- Unfriendly

religion_means <- lgbt1 |>
  summarize(Evangelical_Mean = mean(Evangelical),
            Catholic_Mean = mean(Catholic),
            Jewish_Mean = mean(Jewish),
            Muslim_Mean = mean(Muslim),
            Mormon_Mean = mean(Mormon),
            NonEvangelical_Mean = mean(`Non-Evangelical Protestant`)) |>
  knitr::kable(digits = 3)
  

religion_means
```
When respondents were asked what they assumed numerous religions felt about supporting the LGBT community, they responded the displayed results. It's important to note that there may be some confounding bias in what respondents personally identify as, what religions they are frequently exposed to, or what specific interactions they've had with different religious groups. 
The data showcases that, on the scale of 1-3, where 3 is the most unfriendly, Muslims were perceived to be the least friendly to the LGBT community. The Jewish community was seen to be the most friendly with a score of 2.299. The difference in means between the Muslim and Jewish groups was 0.59, so the variation isn't very large. Another takeaway from this visualization is that all the means from the respondents on these religions shifted towards the unfriendly side of the spectrum. No religion was identified as friendly towards the LGBT community, let alone neutral.



```{r}

imp_binary <- lgbt1 |>
  filter(!Q49 == -1) |>
  mutate(Religious = if_else((IMP == 1 | IMP == 2),
                              "Important",
                              "Not Important"))

## For people who consider religion important in their lives, how many have come out to most important people in their life?

imp_binary |>
  group_by(Religious, Important_People) |>
  summarize(n = n()) |>
  pivot_wider(names_from = Important_People,
              values_from = n) |>
  knitr::kable()

## Things to fix: how to make each row a proportion of itself, change column names to be in specific order (not alphabetical)
```
With this table, I hoped to uncover the discrepancy between religious vs nonreligious LGBT respondents coming out to important people in their lives. One observation is that there are more LGBT individuals who don't consider religion important to them. Among those individuals, most have come out to all/a majority of important people in their lives. Among relgious LGBT respondents, most have also come out to all/a majority of important people in their lives, however a greater proportion say that they have come out to none of the important people in their life, relative to those who are not religious.
There does not seem to be a huge impact of considering religion important to you, on coming out to important people in your life. Yet still, it's interesting that these same respondents who have come out, still identified all relgions are generally unfriendly towards the LGBT. A further data point that would be useful to explore is the reaction of the people that the respondents came out to, as well as what their religious views were. 
